---
title: "USA Mass Shootings in 2015"
output: html_document
---
```{r, echo=FALSE, message=FALSE}
#requre packages
library(jsonlite)
library(dplyr)
library(knitr)
library(plotly)

#Set working directory
#setwd("/Users/Potato/Documents/Portfolio")
#read in data
df <- read.csv("data/Mass-Shooting-Data.csv")
#add a column to show what will be displayed when mouse hovers over data point
df$total <- df$injured + df$killed
df$hover <- paste0(df$city, ", ", df$state, ": ", df$injured, " injured, ", df$killed, " killed.")

#create a table that summarizes data
summary_data <- df %>%summarise(
  total_incidents = n(),
  total_killed = sum(killed),
  avg_killed = round(mean(killed), 2),
  total_injuries = sum(injured),
  avg_injured = round(mean(injured), 2),
  total_casualties = sum(total),
  avg_casualties = round(mean(total), 2)
)

#make table names readable
names(summary_data) <- c("Total Incidents", "Total Killed", "Average Killed", "Total Injured", "Average Injured", 
                         "Total Casualties", "Average Casualties")
summary_table <- kable(summary_data, format = "markdown")

#Find the incident which has most killed/most casualties.
most_killed <- filter(df, killed == max(killed))
most_casualties <- filter(df, total == max(total))

#summarise data by state
state_data <- df %>% group_by(state) %>%
  summarise(
    total_incidents = n(),
    total_casualties = sum(total),
    total_killed = sum(killed),
    mean_casualties = round(mean(total), 2),
    mean_killed = round(mean(killed), 2)
  )

#find noteworthy states using calculated data
highest_incident_state <- filter(state_data, total_incidents == max(total_incidents))
most_casualties_state <- filter(state_data, total_casualties == max(total_casualties))
most_dangerous_state <- filter(state_data, mean_casualties == max(mean_casualties))
most_deadly_state <- filter(state_data, mean_killed == max(mean_killed))


names(state_data) <- c("State", "Total Incidents", "Total Casualties", "Total Killed", "Mean Casualties", "Mean Killed")
state_table <- kable(state_data, format = "markdown")



  
```
In 2015, there were `r summary_data$'Total Incidents'` _total shooting incidents_, which resulted in `r summary_data$'Total Injured'` injuries and `r summary_data$'Total Killed'` deaths. An overall total of `r summary_data$'Total Casualties'` casualties. The single shooting with the _most total casualties_ was in `r most_casualties$city`, `r most_casualties$state` on `r most_casualties$date`, with `r most_casualties$injured` injured and `r most_casualties$killed` dead. There were `r most_casualties$total` casualties. 

### Summary of Shooting Data

`r summary_table`

## State Data

The highest incident state was `r highest_incident_state$state`, which had `r highest_incident_state$total_incidents` incidents in 2015. The state with the most total casualties was `r most_casualties_state$state` which had `r most_casualties_state$total_casualties` casualties in 2015. The state which suffered the most casualties per incident was `r most_dangerous_state$state`, which had an average of `r most_dangerous_state$mean_casualties` casualties per incident. The state that had the most deaths per incident was `r most_deadly_state$state`, which had an average of `r most_deadly_state$mean_killed` deaths per incident.

`r state_table`

## Interactive Data

Below is an interactive map, and an interactive scatterplot, both which include every incident. Roll over an incident with the cursor for more information. 


```{r, echo=FALSE, message=FALSE}

#layout parameters for map
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('grey75'),
  showland = TRUE,
  landcolor = toRGB("grey95"),
  subunitwidth = 1,
  countrywidth = 1,
  subunitcolor = toRGB("gray75"),
  countrycolor = toRGB("gray50")
  
)
#plot map
plot_ly(df, lon = lng, lat = lat, text = hover,
        marker = list(size = total * 1.5, cmin = 0, cmax = 10, symbol = "circle-x"), 
        type = 'scattergeo', 
        locationmode = 'USA-states', 
        color = killed,
        colors = 'Reds',
        opacity = total,
        line = list(color = toRGB("red4"), width = 0.5)
        ) %>%
  layout(title = '2015 US city mass shootings', geo = g)

#plot scatterplot: date vs casualties
plot_ly(df, x = df$date, y = df$total,
        marker = list(cmin = 0, cmax = 10, symbol = "circle-x"),
        text = hover,
        mode = "markers", 
        name = "Casualties Over Time", 
        color = killed, 
        size = killed,
        colors = "Reds",
        opacity = total,
        line = list(color = toRGB("red4"), width = 0.5)
        ) %>%
  layout(xaxis = list(title = "Date"), yaxis = list(title = "Casualties"), hovermode = "closest")
```

